---
title: "Creat phyloseq object"
author: "Ran An"
date: "`r Sys.Date()`"
output: bookdown::gitbook
site: bookdown::bookdown_site
---
#Phyloseq object

**Create Folders as following** 
```{r, eval=FALSE}
#Tables  
dir.create("tables")
# Figures 
dir.create("figures")  
# Phyloseq objects  
dir.create("phyobjects")  
#input files
dir.create("input files")
```

**Load packages**  
```{r, warning=FALSE, message=FALSE}
library(microbiome)
library(phyloseq) 
library(microbiomeutilities) 
library(RColorBrewer) 
library(ggpubr) 
library(DT) 
library(data.table)
library(dplyr) 
```


**Create the phyloseq object**
```{r, warning=FALSE, message=FALSE}
#load the biome and mapping file
ps<- read_phyloseq(otu.file = "./input files/biom_GH002.biom", 
                    taxonomy.file = NULL, 
                    metadata.file = "./input files/mapping_GH002_git.csv", 
                    type = "biom") 
#load tree file
library(ape)
treefile_p1 <- read.tree("./input files/tree_GH002.tree")

#merge the phyloseq object
ps<-merge_phyloseq(ps,treefile_p1)
```

**Clean the taxonomy table**
```{r}
# We make a data table with information on the OTUs
#datatable(tax_table(ps))
tax_table(ps)[,colnames(tax_table(ps))] <- gsub(tax_table(ps)[,colnames(tax_table(ps))],pattern="[a-z]__",replacement="")
tax_table(ps)[tax_table(ps)[,"Phylum"]== "","Phylum"] <- "Unidentified"
datatable(tax_table(ps))#compare the new table with the old table
```

**Save the complete phyloseq object**
```{r}
saveRDS(ps, "./phyobjects/ps_GH002.rds")
```


**Subset the biological samples (i.e.remove MOCK communities)**  
Mock communities are synthesised microbial community which were used to check the senquencing error.Save the phyloseq object after removing mocks - "ps_GH002_rm"
```{r}
library(phyloseq)
ps_GH002_rm<-subset_samples(ps,group != "MOCK")
ps_GH002_rm<-prune_taxa(taxa_sums(ps_GH002_rm) > 0, ps_GH002_rm)
saveRDS(ps_GH002_rm, "./phyobjects/ps_GH002_rm.rds")
ps_GH002_rm
```

Save the phyloseq object of mocks - "ps_GH002_mc"
```{r}
library(phyloseq)
ps_GH002_mc<-subset_samples(ps,group == "MOCK")
ps_GH002_mc<-prune_taxa(taxa_sums(ps_GH002_mc) > 0,ps_GH002_mc)
saveRDS(ps_GH002_mc, "./phyobjects/ps_GH002_mc.rds")
ps_GH002_mc
```

Save the phyloseq object before intervention - "ps_t0" 
```{r}
ps_t0<- subset_samples(ps, Time=="pre")
ps_t0<-prune_taxa(taxa_sums(ps_t0) > 0,ps_t0)
saveRDS(ps_t0, "./phyobjects/ps_t0.rds")
ps_t0
```

Save the phyloseq object after intervention- "ps_t1"
```{r}
ps_t1<- subset_samples(ps, Time=="post")
ps_t1<-prune_taxa(taxa_sums(ps_t1) > 0,ps_t1)
saveRDS(ps_t1, "./phyobjects/ps_t1.rds")
ps_t1
```

```{r}
sessionInfo()
```

