---
title: "Data exploration"
author: "Ran An"
date: "`r Sys.Date()`"
output: bookdown::gitbook
site: bookdown::bookdown_site
---

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(microbiome) # data analysis and visualisation
library(phyloseq) # also the basis of data object. Data analysis and visualisation
library(microbiomeutilities) # some utility tools 
library(RColorBrewer) # nice color options
library(ggpubr) # publication quality figures, based on ggplot2
library(DT) # interactive tables in html and markdown
library(data.table) # alternative to data.frame
library(dplyr) # data handling  
```

##Data exploration

**load pre-stored phyloseq object**
```{r}
#load pre-created phyloseq object
ps_rm <- readRDS("./phyobjects/ps_GH002_rm.rds")
# check Nr. OTU etc.
print(ps_rm)
```

```{r,warning=FALSE, message=FALSE}
#global check up
library(microbiome)
min(sample_sums(ps_rm))
max(sample_sums(ps_rm))
mean(sample_sums(ps_rm))
#summarize_phyloseq(ps_rm)
```

Let us check for distribution of number of sequences retained after NG-tax approach.  
```{r,warning=FALSE, message=FALSE}
#reads distribution
p_seqdepth<- plot_read_distribution(ps_rm, "Group", "density")
print(p_seqdepth)
```


Moving on to distribution of taxa
```{r}
#plot the histogram of ASVs
ps<-ps_rm
library(data.table)
ps0_df_taxa <- data.table(tax_table(ps), 
                        ASVabundance = taxa_sums(ps), 
                        ASV= taxa_names(ps))
ps0_tax_plot <- ggplot(ps0_df_taxa, aes(ASVabundance)) + 
  geom_histogram() + ggtitle("Histogram of ASVs (unique sequence) counts (NG-Tax)") + 
  theme_bw() + scale_x_log10() + ylab("Frequency of ASVs") + xlab("Abundance (raw counts)") + scale_x_continuous(limits = c(0, 500))

print(ps0_tax_plot)
```


**Check how different phyla are represented in the total data**
```{r}
taxasums = rowSums(otu_table(ps))
  
taxatable <- as.data.frame.matrix(tax_table(ps))

tax_plot1 <- ggplot(taxatable, aes(x = taxasums, color = taxatable[, "Phylum"])) 
tax_plot1 <- tax_plot1 + geom_line(size = 1.5, stat = "density") 
tax_plot1 <- tax_plot1 + theme_bw() 
tax_plot1 <- tax_plot1 + guides(color=guide_legend(title="Phylum", nrow = 8)) 
tax_plot1 + scale_x_log10() + xlab("Log10 OTU Counts") 
```


Check out the phylum level (average abundance vs prevalence)
```{r}
p <- plot_taxa_prevalence(ps, "Phylum")
p
```

Checking for potentially spurious OTUs/ASVs. Usually, we do not expect mitochondria and chloroplast as part of the human microbiome.  

ASV-picking stratergy  
```{r}
ps1 <- subset_taxa(ps, Family != "chloroplast")

print(ps1)

# also check how many are lost 

ntaxa(ps1)-ntaxa(ps)
```
There is no mitochrondria and chloroplast in our dataset

